### 2015-05-20 public.tex

\chapter*{在学期间论文发表情况}
\addcontentsline{toc}{chapter}{在学期间论文发表情况}

1. Jianmin Ji, Hai Wan, Ziwei Huo, Zhenfeng Yuan, Simplifying a Logic Program Using Its Consequences, In Proceedings of the Twenty-fifth International Joint Conference on Artificial Intelligence (IJCAI-15). 2015. July 25-31, 2015, Buenos Aires, Argentina.

此论文的内容由本文第四章的工作扩展而来。
\\

%1、Jianmin Ji, Hai Wan, Ziwei Huo, Zhenfeng Yuan. Splitting a Logic Program Revisited. Proceedings of the Twenty-ninth AAAI Conference on Artificial Intelligence (AAAI-15). 2015.
2. Jianmin Ji, Hai Wan, Ziwei Huo, Zhenfeng Yuan, Splitting a Logic Program Revisited, In Proceedings of the Twenty-ninth AAAI Conference on Artificial Intelligence (AAAI-15). 2015. 
January 25–30, 2015, Austin Texas, USA.

本文第三章和第四章中的工作内容为此论文的主要内容。
\\

%2、Jianming Ji, Hai Wan, Peng Xiao, Ziwei Huo and Zhanhao Xiao.Elementary Loop revisited. Proceedings of the Twenty-eighth AAAI Conference on Artificial Intelligence (AAAI-14). 2014.
3. Jianmin Ji, Hai Wan, Peng Xiao, Ziwei Huo, Zhanhao Xiao, Elementary Loops Revisited, In Proceedings of the Twenty-eighth AAAI Conference on Artificial Intelligence (AAAI-14). 2014.
July 27 -31, 2014, Qu{\'{e}}bec City, Qu{\'{e}}bec, Canada.

本文第三章中提出关键半环集的思想基于此论文得到。